---
title: "Data Engineering Workflow"
subtitle: "From Raw Data to Visualization-Ready Dataset"
author:
  - name: Kathy Mok Kai Ting (2402530)
  - name: Siew Yong Ying (2202450)
  - name: Ellie Josepha Lim Shi Ern (2401756)
  - name: Leow Yi Hao Ignatius (2501538)
  - name: Poh Yang Jie Michael (2500624)
date: 2026-03-23
date-format: "dddd MMM DD, YYYY"
execute:
  echo: true
  warning: false
  message: false

format: 
  html:
    toc: true
    toc-location: left
    number-sections: true
    embed-resources: true
    theme: cosmo
    title-block-banner: true
    code-fold: true
    code-summary: "Show/Hide code"
---
# Data Sources

This project uses two primary data sources:

1.  **Eurostat `nrg_ti_gas`** — Annual natural gas trade data by partner country, providing import volumes in million cubic metres (MCM). Downloaded from the [Eurostat Data Browser](https://ec.europa.eu/eurostat/databrowser/view/nrg_ti_gas/default/table?lang=en).

2.  **Ukraine-Russia War Graphics Dataset** — A curated Google Sheet by *Information is Beautiful* containing country-level data on gas dependency percentages, share of Russian exports, and geopolitical stance. Available at [informationisbeautiful.net](https://informationisbeautiful.net/visualizations/ukraine-russia-war-infographics-data-visuals/).

# Step 1 — Load Packages and Read Raw Data

```{r}
#| label: load-packages
#| message: false
#| warning: false

library(readr)
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggrepel)
```

```{r}
#| label: read-eurostat
#| message: false

eurostat_raw <- read_csv(
  "data/nrg_ti_gas__custom_20384524_linear.csv",
  show_col_types = FALSE
)

glimpse(eurostat_raw)
```

```{r}
#| label: read-war-graphics

war_dep <- read_xlsx("data/Ukraine-Russia_War_Graphics.xlsx",
                     sheet = "% dependency on Russian Gas - V")
war_support <- read_xlsx("data/Ukraine-Russia_War_Graphics.xlsx",
                         sheet = "Who Supports Russia")
```

# Step 2 — Clean Eurostat Data

The Eurostat dataset contains annual natural gas imports from Russia and total imports, reported in million cubic metres (MCM). We filter to **2021** as the reference year (the last full year before the invasion) and separate Russian imports from total imports.

```{r}
#| label: clean-eurostat

# Remove aggregate regions (EU, Euro area)
aggregate_geos <- c(
  "European Union - 27 countries (from 2020)",
  "Euro area – 20 countries (2023-2025)",
  "Euro area – 21 countries (from 2026)"
)

eurostat_clean <- eurostat_raw |>
  filter(
    TIME_PERIOD == 2021,
    !(geo %in% aggregate_geos)
  ) |>
  select(geo, partner, OBS_VALUE) |>
  pivot_wider(
    names_from = partner,
    values_from = OBS_VALUE,
    values_fill = 0
  )

# Rename columns for clarity
names(eurostat_clean) <- c("country", "imports_russia_mcm", "imports_total_mcm")

# Convert MCM to BCM (billion cubic metres)
eurostat_clean <- eurostat_clean |>
  mutate(
    imports_russia_bcm = imports_russia_mcm / 1000,
    imports_total_bcm  = imports_total_mcm / 1000
  )

head(eurostat_clean, 10)
```

# Step 3 — Clean War Graphics Data

The War Graphics sheet provides the dependency percentage and share of Russian exports per country. We standardise column names and handle the dual LOW/HIGH estimate entries for Hungary, Austria, and Romania.

```{r}
#| label: clean-war-dep

# Rename columns from the "% dependency on Russian Gas - V" sheet
war_dep_clean <- war_dep |>
  select(
    country    = 1,
    note       = 2,
    dep_pct    = 3,
    export_pct = 4
  ) |>
  filter(!is.na(country)) |>
  mutate(
    dep_pct    = as.numeric(dep_pct),
    export_pct = as.numeric(export_pct)
  )

# --- Improvement 3: Consolidate uncertainty ranges ---
# For countries with dual estimates (upper/lower), compute a single midpoint
# and flag them as uncertain entries
dual_entries <- war_dep_clean |>
  filter(str_detect(note, "estimate", negate = FALSE)) |>
  group_by(country) |>
  summarise(
    dep_pct    = mean(dep_pct, na.rm = TRUE),
    export_pct = first(export_pct),
    is_estimate = TRUE,
    .groups = "drop"
  )

single_entries <- war_dep_clean |>
  filter(is.na(note) | !str_detect(note, "estimate")) |>
  mutate(is_estimate = FALSE) |>
  select(country, dep_pct, export_pct, is_estimate)

war_dep_final <- bind_rows(single_entries, dual_entries)

head(war_dep_final, 10)
```

# Step 4 — Clean Geopolitical Alignment Data

We simplify the five-level geopolitical response into three categories for the colour aesthetic.

```{r}
#| label: clean-support

war_support_clean <- war_support |>
  select(country = 1, response = 2) |>
  filter(!is.na(country), !is.na(response), response != "") |>
  mutate(
    geopolitical_alignment = case_when(
      response %in% c("PRO RUSSIA", "LEANS RUSSIA") ~ "Pro-Russia / Leaning",
      response %in% c("CONDEMNS RUSSIA")             ~ "Opposes Russia",
      response %in% c("NEUTRAL", "UNHAPPY")           ~ "Neutral / Non-aligned",
      TRUE                                             ~ NA_character_
    )
  ) |>
  filter(!is.na(geopolitical_alignment)) |>
  select(country, geopolitical_alignment)

# Standardise country names to match across datasets
war_support_clean <- war_support_clean |>
  mutate(country = case_when(
    country == "Bosnia & Herzegovina" ~ "Bosnia",
    country == "South Korea"          ~ "South Korea",
    TRUE                              ~ country
  ))

head(war_support_clean, 10)
```

# Step 5 — Merge Datasets

We merge the three cleaned datasets using `left_join()` on country name.

```{r}
#| label: merge-data

# Start from the war dependency data (most complete for our viz)
viz_data <- war_dep_final |>
  left_join(
    eurostat_clean |> select(country, imports_russia_bcm),
    by = "country"
  ) |>
  left_join(
    war_support_clean,
    by = "country"
  )

# Standardise Eurostat names to match war graphics names
# Some countries need manual mapping
name_map <- c(
  "Türkiye"          = "Turkey",
  "Czechia"          = "Czechia",
  "North Macedonia"  = "North Macedonia",
  "Bosnia and Herzegovina" = "Bosnia"
)

# Re-do the Eurostat join with name harmonisation
eurostat_renamed <- eurostat_clean |>
  mutate(country = ifelse(country %in% names(name_map),
                           name_map[country], country))

viz_data <- war_dep_final |>
  left_join(
    eurostat_renamed |> select(country, imports_russia_bcm),
    by = "country"
  ) |>
  left_join(
    war_support_clean,
    by = "country"
  )

# Fill missing alignment with "Unknown"
viz_data <- viz_data |>
  mutate(
    geopolitical_alignment = replace_na(geopolitical_alignment, "Unknown"),
    # Convert export_pct from proportion to percentage for display
    export_pct_display = export_pct * 100,
    # Convert dep_pct from proportion to percentage for display
    dep_pct_display = dep_pct * 100
  )

# Remove rows with no dependency data
viz_data <- viz_data |>
  filter(!is.na(dep_pct))

glimpse(viz_data)
```

# Step 6 — Final Visualization-Ready Dataset

```{r}
#| label: final-dataset

# Make geopolitical_alignment an ordered factor for consistent legend ordering
viz_data <- viz_data |>
  mutate(
    geopolitical_alignment = factor(
      geopolitical_alignment,
      levels = c("Opposes Russia", "Neutral / Non-aligned",
                 "Pro-Russia / Leaning", "Unknown")
    )
  )

# Display final dataset
viz_data |>
  arrange(desc(dep_pct)) |>
  select(country, dep_pct_display, export_pct_display,
         imports_russia_bcm, geopolitical_alignment, is_estimate) |>
  head(20)
```

```{r}
#| label: save-viz-data

write_csv(viz_data, "data/viz_ready.csv")
```

The visualization-ready dataset contains **`r nrow(viz_data)` countries** with the following variables:

- `country` — Country name
- `dep_pct` / `dep_pct_display` — Dependency on Russian gas (proportion / percentage)
- `export_pct` / `export_pct_display` — Share of Russia's total gas exports (proportion / percentage)
- `imports_russia_bcm` — Absolute gas imports from Russia in billion cubic metres (2021)
- `geopolitical_alignment` — Categorical: Opposes Russia, Neutral / Non-aligned, Pro-Russia / Leaning
- `is_estimate` — Logical flag for countries with consolidated uncertainty ranges
